---
title: "FLS2"
author: "Hayoung Cheon"
date: "2024-09-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r}
#ggplot2
library(ggplot2)
library(plotly)
library(ggthemes)

```

## Import PlayerBBall.csv
```{r}
players = read.csv("/Users/oyunm/Downloads/PlayersBBall.csv", header = TRUE)
#players

#players$name
#players$year_start
#players$year_end
#players$position
#players$height
#players$weight
#players$birth_date

```
### Clean the data
```{r}
#Identify the missing data in position
which(players$position=="") #2143
players[2143,]            # his position height and weight is missing

players[2143,]$position = "G"
players[2143,]$height = "6-2"
players[2143,]$weight= "185"

players[2143,]

## change C-F to F-C
players$position <- gsub("C-F", "F-C", players$position)

## change G-F to F-G
players$position <- gsub("G-F", "F-G", players$position)

#To check if replacing has gone right.
players[2141,]
```
### number of players by position
```{r}
players %>%
  ggplot(aes(x=position, fill=position)) + geom_bar() +
  ggtitle("Number of Basketball Players by position") +
  theme_excel()

```


### weight of center vs. forwards
```{r}
# since the weight data was in character thhe box plot wasn't showing correctly. Changed the data into numeric.
players$weight <- as.numeric(players$weight)

players %>% filter(players$position == "C" | players$position == "F") %>%
  ggplot(aes(x=position, y=weight, fill=position)) + geom_boxplot() +
  ggtitle("Weight of Center vs. Forwards") 


```

### height of center vs. forwards
```{r}
#factor() function helps to change the order of the axis columns
summary(players$height)
players$height_change = factor(players$height, levels = c("5-0","5-1", "5-2", "5-3","5-4","5-5","5-6","5-7","5-8","5-9","5-10","5-11","5-12","6-0", "6-1", "6-2", "6-3", "6-4", "6-5", "6-6", "6-7", "6-8", "6-9", "6-10", "6-11", "6-12", "7-1", "7-2", "7-3", "7-4", "7-5", "7-6", "7-7"))
n <- is.na(players$height)
players %>% filter(players$position == "C" | players$position == "F") %>%
  ggplot(aes(x=height_change, fill=position)) + geom_bar() +
  ggtitle("Height of Center vs. Forwards")

```

### height difference between positions
```{r}
players %>%
  ggplot(aes(x=height_change, fill=position)) +geom_bar() + 
  ggtitle("Height distribution by positions") +xlab("height")+
  facet_wrap(~position)
```

## Change height(Character type) data type to a numeric type
```{r}
library("tidyr")

#Change heights character type to numeric
players <- players %>%
separate(height, into = c("feet", "inches"), sep = "-", convert = TRUE)
players$height_inches <- players$feet * 12 + players$inches
```

### A relation between height and weight
```{r}
ggplot(data = players, mapping = aes(x=weight, y=height_inches)) +
  geom_point() + geom_smooth(method = "lm")+
  ggtitle("Weight vs. Height") + ylab("height")

```

### A relation between height and weight by position
```{r}

ggplot(data = players, mapping = aes(x = weight, y=height_inches, linetype=position, color=position)) +
  geom_point() +geom_smooth(method = "lm", fill="black") +
  facet_wrap(~position)
```

### Height increase over years
```{r}
ggplot(data=players, mapping = aes(x= year_start, y=height_inches)) +
  geom_point() +geom_smooth(method="lm") +
  ggtitle("Height vs. Year Started") + xlab("Year Started")+ylab("Height")

```

### 3D plots of weight vs. height vs. years
```{r}
p <- plot_ly(players, x=~weight, y=~height_inches, z=~year_start, color = ~position) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = "Weight"),
                      yaxis = list(title = "Height"),
                      zaxis = list(title = "Year Started")))

ggplotly(p)
```